---
title: "ASV differences between wild and F2"
author: "Lei Ma"
output: github_document
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r libraries, echo=FALSE, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, include=TRUE)
knitr::knit_hooks$set(inline = function(x) if(is.numeric(x)){format(x,digits=2)}else{x})

#misc
library(here)

#data manipulations
library(tidyverse)
library(readr)

#plotting stuff
library(ggplot2)

#analysis
library(phyloseq)
library(vegan)
library(VennDiagram)
library(corncob)
```

```{r data}
ps <- readRDS(here("data/processed/ps.rds"))
ps.rel <- readRDS(here("data/processed/psrel.rds"))
ps.rel.melted <- readRDS(here("data/processed/psrelmelted.rds"))
```

## Overlap of ASVs

I was curious how many ASVs were unique to each environment/fish type: Scorton Creek wild, New Bedford Harbor wild, and the respective F2 fish

```{r venn}
taxa_list <- list(sc_w = subset_samples(ps, fishType=="Scorton Creek wild") %>% prune_taxa(taxa_sums(.)>0,.) %>% taxa_names(),
                  nb_w = subset_samples(ps, fishType=="New Bedford Harbor wild") %>% prune_taxa(taxa_sums(.)>0,.) %>% taxa_names(),
                  sc_f2 = subset_samples(ps, fishType=="Scorton Creek F2") %>% prune_taxa(taxa_sums(.)>0,.) %>% taxa_names(),
                  nb_f2 = subset_samples(ps, fishType=="New Bedford Harbor F2") %>% prune_taxa(taxa_sums(.)>0,.) %>% taxa_names())

taxa_overlap <- calculate.overlap(taxa_list)

venn.diagram(taxa_list, filename = here("figures/Venn_gut.tiff"), imagetype="tiff", main="Venn diagram of taxa shared between fish types")
knitr::include_graphics(here("figures/Venn_gut.tiff"))
```
## "Core taxa" present in all fish types

Core taxa in wild type are either Vibrionaceae or Mycoplasma while core taxa in F2 are Enterococcus
```{r core}
ps.core <- prune_taxa(taxa_names(ps.rel) %in% taxa_overlap$a6, ps.rel) %>% subset_samples(sampleType=="gut")
core.plot <- ps.core %>% plot_bar(fill="Order")+ geom_bar(aes(fill=Order), stat="identity", position="stack")+facet_wrap(~fishType, scales="free_x")
core.plot
```

Are the vibrios/mycoplasmatales and enterococcus negative associated with each other within wild types and F2s? Or is it just that enterococcus overtook the vibrios? Analysis TBD. Maybe with heatmaps?

```{r heatmap}

```